When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-03-01 paper/paper.tex
Enn Pee
More Enn Pee

\begin{abstract}
    The subgraph isomorphism problem involves finding a copy of a pattern graph inside a larger
    target graph. The non-induced version allows extra edges in the target, whilst the induced
    version does not.  Although both variants are \NP-complete, algorithms inspired by constraint programming
    can operate comfortably on many real-world problem instances with thousands of vertices.
    However, they cannot handle arbitrary instances of this size. We show how to generate ``really
    hard'' random instances for subgraph isomorphism problems, which are computationally challenging
    with a couple of hundred vertices in the target, and only twenty pattern vertices. For the
    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
    transition, whose location we can predict; for the induced variant, much richer behaviour is
    observed, and constrainedness gives a better measure of difficulty than does proximity to a
    phase transition.  These results have practical consequences: we explain why the widely
    researched ``filter~/ verify'' indexing technique used in graph databases is founded upon a
    misunderstanding of the empirical hardness of \NP-complete problems, and cannot be beneficial
    when paired with any reasonable subgraph isomorphism algorithm.
\end{abstract}

@@

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent vertices $v$ and $w$ in
$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. Where labels are present,
the mapping must also preserve labels, so $\ell(v) = \ell(i(v))$ for any $v$. An \emph{induced
subgraph isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not
adjacent in $P$, then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P
\rightarrowtail T$ for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced
isomorphism. Observe that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced
isomorphism $i : P \rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} :
\shortoverline{P} \rightarrowtail \shortoverline{T}$. Finding either form of isomorphism between two
given graphs is \NP-complete.

>>>>

\begin{abstract}
    The subgraph isomorphism problem involves deciding whether a copy of a pattern graph inside a larger
    target graph. The non-induced version allows extra edges in the target, whilst the induced
    version does not.  Although both variants are \NP-complete, algorithms inspired by constraint programming
    can operate comfortably on many real-world problem instances with thousands of vertices.
    However, they cannot handle arbitrary instances of this size. We show how to generate ``really
    hard'' random instances for subgraph isomorphism problems, which are computationally challenging
    with a couple of hundred vertices in the target, and only twenty pattern vertices. For the
    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
    transition, whose location we can predict; for the induced variant, much richer behaviour is
    observed, and constrainedness gives a better measure of difficulty than does proximity to a
    phase transition.  These results have practical consequences: we explain why the widely
    researched ``filter~/ verify'' indexing technique used in graph databases is founded upon a
    misunderstanding of the empirical hardness of \NP-complete problems, and cannot be beneficial
    when paired with any reasonable subgraph isomorphism algorithm.
\end{abstract}

@@

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent vertices $v$ and $w$ in
$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. Where labels are present,
the mapping must also preserve labels, so $\ell(v) = \ell(i(v))$ for any $v$. An \emph{induced
subgraph isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not
adjacent in $P$, then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P
\rightarrowtail T$ for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced
isomorphism. Observe that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced
isomorphism $i : P \rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} :
\shortoverline{P} \rightarrowtail \shortoverline{T}$. Deciding whether a subisomorphism of either
form exists between two given graphs exists is \NP-complete.
